`r if(knitr:::is_latex_output()) '\\appendix'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'` 
<!--
The first appendix must start with the above text. Do not remove!
-->

# Main Appendix

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readability and/or setup.

## In [Chapter 1: Introduction](#intro)

No noteworthy code for this chapter.

## In [Chapter 2: Simulations](#simulations)

Code for the reorganization of dataframes in order to create tables.

```{r app_tablesorg, ref.label="tablesorg", echo=TRUE, eval=FALSE}
```

```{r app_lntable, ref.label="lntable", echo=TRUE, eval=FALSE}
```

```{r app_exptable, ref.label="exptable", echo=TRUE, eval=FALSE}
```

```{r app_wtable, ref.label="wtable", echo=TRUE, eval=FALSE}
```

## In [Chapter 3: Case Study](#casestudy)

### Data Dictionary

Code for the creation of the data dictionary for our dataset.

```{r app_datadict, ref.label="datadict", echo=TRUE, eval=FALSE}
```

### Map

Code for the creation of the US map visualization.

```{r app_usmap, ref.label="usmap", echo=TRUE, eval=FALSE}
```

### Health-based Thresholds

Code for the creation of the health-based thresholds dataframe.

```{r app_hbttable, ref.label="hbttable", echo=TRUE, eval=FALSE}
```

### Top 10 Tables

Code for the creation of our tables for the top 10 sites, according to a specified method.

```{r app_top10baseline, ref.label="top10baseline", echo=TRUE, eval=FALSE}
```

```{r app_top10substitution, ref.label="top10substitution", echo=TRUE, eval=FALSE}
```

```{r app_top10km, ref.label="top10km", echo=TRUE, eval=FALSE}
```

## In [Chapter 4: Conclusion](#conclusion)

No noteworthy code for this chapter.
