---
title: "Coal Ash Data"
author: ""
date: ""
output: 
  pdf_document:
    fig_width: 3.5
    
---

```{r, setup, include=FALSE}
# load packages
require(tidyverse)
require(mosaic) 
require(kableExtra)
require(forcats)
require(GGally)
require(plotly)
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small",
  warning = FALSE,
  message = FALSE,
  echo = FALSE, 
  cache = TRUE)   # slightly smaller font for code
theme_set(theme_classic()) # set ggplot theme
figdir <- "figures/"
```


# Read in results and identify exclusion criteria
FIELD	DESCRIPTION

* `state`	The state where the site is located.
* `site`	The name of the site as it is presented in its groundwater monitoring report.
* `disposal.area`	The name of the disposal area(s) as they are presented in the groundwater monitoring report. Note: some wells (`well.id`) monitor groundwater from more than one disposal unit.
* `type`	The type of disposal unit. SI = surface impoundment, L= landfill, M = mixed multi-unit (landfill and surface impoundment), and U = unknown.
* `well.id`	The identifier given to each monitoring well in the groundwater monitoring report.
* `gradient`	The location of the groundwater monitoring well relative to the regulated ash disposal unit it monitors. (may be up/downgradient from multiple disposal areas.)
* `sample.date`	The date the well was sampled.
* `contaminant`	The contaminant name. These have been standardized to allow for analyses across plants.
* `measurement.unit`	The concentration units. These include mg/l, ug/l, pCi/l, and standard units (SU) for pH.
* `below.detection`	"<" Indicates that a contaminant was not detected. In this case, the value in the "concentration" column will be the method detection limit (if known) or the reporting limit. 
* `concentration`	The monitoring result (or, for nondetects, the method detection limit or reporting limit).
qualifier	Lab qualifiers assigned to the result. See each monitoring report for detailed explanations.
link	The link to the groundwater monitoring report.

\footnotesize
```{r readdata}
# takes about 10 minutes to run
# resulting dataframes (fulldata, fences) are stored in fulldata.RData
coal <- read_csv("../../21_tony-ni_harvard-summer-biostats/data/chemical_data.csv") %>%
  # ph was listed as mg/l for several cases
  mutate(measurement.unit = if_else(contaminant %in% c("pH", "pH, field"), 
                                    "su", 
                                    measurement.unit))
#,
 #        stdconc.mgl = case_when(measurement.unit == "pCi/l" ~ measurement.unit*0.0015))
```



\newpage
# Detection limits
```{r}
group_by(coal, contaminant, measurement.unit, below.detection) %>%
  summarize(n = n(), 
            lod = mean(concentration), 
            minlod = min(concentration), 
            maxlod = max(concentration), 
            varlod = var(concentration)) %>%
  kable(booktabs = TRUE, digits = 1, 
        longtable = TRUE,
        linesep = c('', '', '', '\\addlinespace')
        )

group_by(coal, measurement.unit, contaminant, below.detection) %>%
  summarize(n = n(), 
            lod = mean(concentration), 
            minlod = min(concentration), 
            maxlod = max(concentration), 
            varlod = var(concentration)) %>%
  kable(booktabs = TRUE, digits = 1, 
        longtable = TRUE,
        linesep = c('', '', '', '\\addlinespace')
        )
```

## Count
```{r below.detection}
pp <- group_by(coal, contaminant, measurement.unit) %>%
  summarize(n = n(), below.detection = sum(!is.na(below.detection))/n) %>%
  arrange(desc(below.detection)) 
kable(pp,booktabs = TRUE, digits = 2, longtable = TRUE)
```

## Proportion below detection
```{r below.detection.prop}
pp %>% arrange(desc(n)) %>%
 kable(booktabs = TRUE, digits = 2, longtable = TRUE)
```

\newpage
# Data structure
```{r, structure}
group_by(coal, state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))

group_by(coal, state, site) %>% 
  summarize(n = n()) %>% 
  arrange(state, desc(n))

group_by(coal, state, site, disposal.area) %>% 
  summarize(n = n()) %>% 
  arrange(state, site, desc(n))
```

# BPA data

```{r}
bpa <- read_delim("bpa_tds2.csv", delim = ";")
bparaw <- read_csv("donnees-bpa-eat2-lhn.csv")
names(bparaw) <- c('foodgroup', 'label', 'type', 
                   'date', 'region', 'vague', 
                   'bpa', 'lod', 'units')
group_by(bparaw, foodgroup, label) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```
