---
title: "simulation-study"
author: "Tony Ni"
date: "1/20/2021"
output: pdf_document
---

```{r}
library(tidyverse)
library(Metrics) #package to help calculate mse
library(NADA) #package with implementation of many methods
library(survival)
```


First, a function to help us generate our desired dataset, where users can specify the mean, sd, sample size (of each run), and censoring rate:

```{r, echo = FALSE}
#function will generate a vector of numbers from the lognormal 
#distribution and censor them at the given rate
#function will take in arguments for 1) samplesize, 2) logmean, 3)logsd
#4) censoring rate

generateLN <- function(sampsize, m, s, censrate){
  true.value <- rlnorm(sampsize, 
               meanlog=log(m^2 / sqrt(s^2 + m^2)),
               sdlog=sqrt(log(1 + (s^2 / m^2))))
  
  uncensored_df <- as.data.frame(true.value) %>%
    arrange(true.value)
  
  censored_df <- uncensored_df %>% #take the head(%) of data to be censored
    slice_head(n=nrow(uncensored_df)*censrate) %>%
    mutate(censored = TRUE)
  
  #full join original df and sliced df
  return_df <- full_join(uncensored_df, censored_df, by = "true.value")

  #replace NAs with FALSE
  return_df$censored <- replace_na(return_df$censored, replace = FALSE)
  
  return(return_df)
}
```

```{r}
#function will generate a vector of numbers from the exponential 
#distribution and censor them at the given rate
#function will take in arguments for 1) samplesize, 2) rate,
#3) censoring rate

generateEXP <- function(sampsize, r, censrate){
  true.value <- rexp(sampsize, rate = r)
  
  uncensored_df <- as.data.frame(true.value) %>%
    arrange(true.value)
  
  censored_df <- uncensored_df %>% #take the head(%) of data to be censored
    slice_head(n=nrow(uncensored_df)*censrate) %>%
    mutate(censored = TRUE)
  
  #full join original df and sliced df
  return_df <- full_join(uncensored_df, censored_df, by = "true.value")

  #replace NAs with FALSE
  return_df$censored <- replace_na(return_df$censored, replace = FALSE)
  
  return(return_df)
}
```

```{r}
#function will generate a vector of numbers from the Weibull
#distribution and censor them at the given rate
#function will take in arguments for 1) samplesize, 2) rate,
#3) censoring rate

generateW <- function(sampsize, sh, sc, censrate){
  true.value <- rweibull(sampsize, shape = sh, scale = sc)
  
  uncensored_df <- as.data.frame(true.value) %>%
    arrange(true.value)
  
  censored_df <- uncensored_df %>% #take the head(%) of data to be censored
    slice_head(n=nrow(uncensored_df)*censrate) %>%
    mutate(censored = TRUE)
  
  #full join original df and sliced df
  return_df <- full_join(uncensored_df, censored_df, by = "true.value")

  #replace NAs with FALSE
  return_df$censored <- replace_na(return_df$censored, replace = FALSE)
  
  return(return_df)
}
```

What do we want our code to do? 

We first need to generate simulated data sets with the following combinations of censoring rates and sample sizes: (censoring rates: 10, 30, 50, sample sizes: 10, 100, 1000)

For each censoring/sample size pair, we create a simulated dataset with those specifications and run the substitution method with a specified number of iterations (`iterations`). For each sample, we will obtain the true mean (when considering all values to be uncensored), the estimated mean (when considering the uncensored values and the imputed values), alongside the MSE of each dataset`       `

the average mse is ultimately what we are interested in

```{r}
iterations <- 1000 #number of iterations
censvalues <- c(0.10, 0.30, 0.50)
sampsizes <- c(10, 100, 1000)

df.tall <- data.frame(prop_cens = numeric(),
                      samplesize = numeric(),
                      iteration = numeric(),
                      method = character(),
                      true_mean = numeric(),
                      mean_complete = numeric(),
                      mean_method = numeric(),
                      true_sd = numeric(),
                      SE_complete = numeric(),
                      SE_method = numeric())
```

#Log-normal

```{r, echo = FALSE, cache = TRUE}
options(scipen=999) #prevent scientific notation
set.seed(7271999)

for(i in censvalues){
  for(j in sampsizes){
    for(k in 1:iterations){
      m <- 1
      s <- 0.5
      df <- generateLN(sampsize = j, m = 1, s = 0.5, censrate = i)
      
      #substitution
      #define LOD to be smallest, uncensored value
      LOD <- min(df$true.value[df$censored == FALSE]) 
      df <- df %>%
        mutate(impSubValue = if_else(censored == TRUE, LOD/2, true.value))
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "substitution",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(df$impSubValue),
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = 
                  sd(df$impSubValue)/sqrt((length(df$impSubValue))))
      
      #mle
      mle_res = cenmle(df$true.value, df$censored)
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "mle",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(mle_res)[1],
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = mean(mle_res)[2])
      
      #km
      km_res = cenfit(df$true.value, df$censored)
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "km",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(km_res)[[1]],
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = mean(km_res)[[2]])
      
      #ros
      ros_res = ros(df$true.value, df$censored)
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "ros",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(ros_res),
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = 
                  sd(ros_res)/sqrt((length(df$true.value))))
    }
    #end of # iterations
  }
}

#aggregating performance criteria

df.ln <- df.tall %>%
  group_by(prop_cens, samplesize, method) %>%
  summarize(Avg_Mean = mean(mean_method),
            Bias = (mean(mean_method) - true_mean),
            Variance = var(mean_method),
            MSE = mse(true_mean, mean_method) 
            ) %>%
  distinct()
```


#Exponential
#not working atm -- having isses with cenmle and specifying exponential distribution...
```{r, echo = FALSE, cache = TRUE, eval = FALSE}
options(scipen=999) #prevent scientific notation
set.seed(7271999)

for(i in censvalues){
  for(j in sampsizes){
    for(k in 1:iterations){
      r = 1
      df <- generateEXP(sampsize = j, r = r, censrate = i)

      #substitution
      #define LOD to be smallest, uncensored value
      LOD <- min(df$true.value[df$censored == FALSE]) 
      df <- df %>%
        mutate(impSubValue = if_else(censored == TRUE, LOD/2, true.value))
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "substitution",
                true_mean = 1/r,
                mean_complete = mean(df$true.value),
                mean_method = mean(df$impSubValue),
                true_sd = 1/r,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = 
                  sd(df$impSubValue)/sqrt((length(df$impSubValue))))
      
      #mle
      mle_res = cenmle(df$true.value, df$censored, dist = "exp")
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "mle",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(mle_res)[1],
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = mean(mle_res)[2])
      
      #km
      km_res = cenfit(df$true.value, df$censored)
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "km",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(km_res)[[1]],
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = mean(km_res)[[2]])
      
      #ros
      ros_res = ros(df$true.value, df$censored)
      
      df.tall <- df.tall %>%
        add_row(prop_cens = i,
                samplesize = j,
                iteration = k,
                method = "ros",
                true_mean = m,
                mean_complete = mean(df$true.value),
                mean_method = mean(ros_res),
                true_sd = s,
                SE_complete = 
                  sd(df$true.value)/sqrt((length(df$true.value))),
                SE_method = 
                  sd(ros_res)/sqrt((length(df$true.value))))
    }
    #end of # iterations
  }
}

#aggregating performance criteria

df.exp <- df.tall %>%
  group_by(prop_cens, samplesize, method) %>%
  summarize(Avg_Mean = mean(mean_method),
            Bias = (mean(mean_method) - true_mean),
            Variance = var(mean_method),
            MSE = mse(true_mean, mean_method) 
            ) %>%
  distinct()
```

```{r}
test <- cenmle(df$true.value, df$censored, dist = 'exponential')
survreg()
survfit(Surv(df$true.value, df$censored))
```

\newpage

```{r}
knitr::kable(df.ln)
```

<!--
#mean by group or disposal area and obtain the mean or something
#graph density distributions from coal and see how theyre distributed (in EVERYTHING facet by state??? site??? something???)
#update methods if time allows
-->
