---
title: "simulation-study"
author: "Tony Ni"
date: "10/8/2020"
output: pdf_document
---

```{r}
library(tidyverse)
library(missMethods) #package for generating missing data
library(NADA) #package for mle for left censored data
library(fitdistrplus) #package for mle imputation method
```

# Preliminary

```{r}
#set seed
set.seed(7271999)

#size of a sample
n.iter <- 100 #number of iterations

#GENERATE POSITIVE VALUES ONLY (MAKE SURE ITS REALISTIC)

mean = 10
sd = 1

#set arbitrary threshold (values below this are considered left-censored)
threshold <- 9

#creating df outside of loop to store summary stats of methods
store_names <- c("iteration", "num_samples", "method", "x_bar", "s^2") #standard variable names -> change s^2 to s.2??
store_df <- as.data.frame(matrix(ncol = 5, nrow = 0))
names(store_df) <- store_names
```

maybe get standard errors in the future (sqrt of those) in the future

# ALL METHODS COMBINED 

```{r}
#writing function to run all methods combined
all_function <- function(n.iter, samplesize){
  for(i in 1:n.iter){
    iteration <- i
    true.value <- rnorm(n = samplesize, mean = mean, sd = sd) #change to true.value
    sim_df <- as.data.frame(cbind(iteration, true.value)) %>%
        #below is T if value below threshold, else F
        mutate(below = as.logical(case_when(true.value <= threshold ~ TRUE, 
                                            true.value > threshold ~ FALSE)),
               #hide values below threshold
               censored.value = if_else(below == FALSE, 
                                        true.value, NA_real_),
               
               #approach #1: substitution
               imp_substitution = if_else(below == TRUE, 
                                          threshold/2, true.value)
        )

    #getting summary stats for substitution
    store_df[i, 1] <- iteration
    store_df[i, 2] <- samplesize
    store_df[i, 3] <- "substitution"
    store_df[i, 4] <- mean(sim_df$imp_substitution)
    store_df[i, 5] <- sd(sim_df$imp_substitution)
    
    #approach #2: mle
    mle_res = cenmle(sim_df$true.value, sim_df$below, 
                     conf.int=0.95, dist = "gaussian")
    
    #getting summary stats for mle 
    store_df[i+n.iter, 1] <- iteration
    store_df[i+n.iter, 2] <- samplesize
    store_df[i+n.iter, 3] <- "mle"
    store_df[i+n.iter, 4] <- coef(summary(mle_res))["(Intercept)"]
    store_df[i+n.iter, 5] <- sd(mle_res)
    
    #approach #3: KM
    km_res = cenfit(sim_df$true.value, sim_df$below, 
                    conf.int=0.95)
    
    #getting summary stats for KM
    store_df[i+(2*n.iter), 1] <- iteration
    store_df[i+(2*n.iter), 2] <- samplesize
    store_df[i+(2*n.iter), 3] <- "km"
    store_df[i+(2*n.iter), 4] <- mean(km_res)[[1]]
    store_df[i+(2*n.iter), 5] <- sd(km_res)
    
    #approach #4 ROS
    ros_res = ros(sim_df$true.value, sim_df$below, 
                  forwardT="log", reverseT="exp")
    
    #getting summary stats for ROS
    store_df[i+(3*n.iter), 1] <- iteration
    store_df[i+(3*n.iter), 2] <- samplesize
    store_df[i+(3*n.iter), 3] <- "ROS"
    store_df[i+(3*n.iter), 4] <- mean(ros_res)
    store_df[i+(3*n.iter), 5] <- sd(ros_res)
    
  }
  return(store_df)
}
```

```{r}
test_all_df <- all_function(10, 100)
```

# new approach (maybe write a function for each method, then run 
# each method with mosaic:do(# here))?

```{r}
sub_func <- function(n, samplesize){
   for(i in 1:n){
    value <- rnorm(n = samplesize, mean = mean, sd = sd)
    id <- seq(1, samplesize, by = 1)
    current_df <- as.data.frame(cbind(id, value))
    
    #creating modified df to have censored info
    modified_df <- current_df %>%
      #below is T if value below threshold, else F
      mutate(below = case_when(value <= threshold ~ "T", 
                               value > threshold ~ "F")
      ) %>%
      #hide values below threshold
      mutate(value = if_else(
        below == "F", value, NA_real_)
      )
    
    #approach #1: substitution
    substitution_df <- modified_df %>%
      mutate(imputed_value = if_else(
        below == "T", 9/2, value)
      )
    
    #getting summary stats for substitution df
    store_df[i, 1] <- paste("samples=", n, ", size=", samplesize, sep = "")
    store_df[i, 2] <- "substitution"
    store_df[i, 3] <- mean(substitution_df$imputed_value)
    store_df[i, 4] <- sd(substitution_df$imputed_value)
  }
  return(store_df)
}

mle_func <- function(n, samplesize){
   for(i in 1:n){
    value <- rnorm(n = samplesize, mean = mean, sd = sd)
    id <- seq(1, samplesize, by = 1)
    current_df <- as.data.frame(cbind(id, value))
    
    #creating modified df to have censored info
    modified_df <- current_df %>%
      #below is T if value below threshold, else F
      mutate(below = case_when(value <= threshold ~ "T", 
                               value > threshold ~ "F")
      ) %>%
      #hide values below threshold
      mutate(value = if_else(
        below == "F", value, NA_real_)
      )
    
    #approach #2: mle
    logical_df <- modified_df %>%
      mutate(below = as.logical(below))

    mle_res = cenmle(logical_df$value, logical_df$below, 
                     conf.int=0.95, dist = "gaussian")
    
    #getting summary stats for mle 
    store_df[i, 1] <- samplesize
    store_df[i, 2] <- "mle"
    store_df[i, 3] <- coef(summary(mle_res))["(Intercept)"]
    store_df[i, 4] <- sd(mle_res)
  }
  return(store_df)
}

ros_func <- function(n, samplesize){ 
   for(i in 1:n){
    value <- rnorm(n = samplesize, mean = mean, sd = sd)
    id <- seq(1, samplesize, by = 1)
    current_df <- as.data.frame(cbind(id, value))
    
    #creating modified df to have censored info
    modified_df <- current_df %>%
      #below is T if value below threshold, else F
      mutate(below = case_when(value <= threshold ~ "T", 
                               value > threshold ~ "F")
      )
    
    #approach #3: ROS
    logical_df <- modified_df %>%
      mutate(below = as.logical(below))

    ros_res = ros(logical_df$value, logical_df$below)
    
    #getting summary stats for km
    store_df[i, 1] <- paste("size=", samplesize, sep = "")
    store_df[i, 2] <- "ros"
    store_df[i, 3] <- mean(ros_res)
    store_df[i, 4] <- sd(ros_res)
    
    NADA::mean(km_res)[[1]]
    NADA::sd(km_res)
    
  }
  return(store_df)
}

km_func <- function(n, samplesize){
  for(i in 1:n){
    value <- rnorm(n = samplesize, mean = mean, sd = sd)
    id <- seq(1, samplesize, by = 1)
    current_df <- as.data.frame(cbind(id, value))
    
    #creating modified df to have censored info
    modified_df <- current_df %>%
      #below is T if value below threshold, else F
      mutate(below = case_when(value <= threshold ~ "T", 
                               value > threshold ~ "F")
      ) %>%
      #hide values below threshold
      mutate(value = if_else(
        below == "F", value, NA_real_)
      )
    
    logical_df <- modified_df %>%
      mutate(below = as.logical(below))
    
    #approach #3: KM
    km_res = cenfit(logical_df$value, logical_df$below, 
                    conf.int=0.95)
    
    #getting summary stats for km
    store_df[i, 1] <- paste("size=", samplesize, sep = "")
    store_df[i, 2] <- "km"
    store_df[i, 3] <- mean(km_res)[[1]]
    store_df[i, 4] <- sd(km_res)
  }
}

mle_imp_func <- function(n, samplesize){
  
}
```

```{r}
testdf1 <- mosaic::do(1) * sub_func(5, 1000)
testdf2 <- mosaic::do(1) * mle_func(5, 1000)
testdf3 <- mosaic::do(1) * ros_func(5, 1000) #errors may arise due to attempts to take negative logs, 
                                             #but can't be helped with out parametric setting...
testdf4 <- mosaic::do(1) * km_func(5, 1000)
```

instead of trying to run the km function, could we try to have 2 separate helper functions? one to pull out the mean and another to pull out the sd? --> NOT SUCCESSFUL

AHHHHHHHH!

hmm... let's just get rid of the KM method fully, i found a new method robust ROS (regression on order statistics) -- which could potentially serve as another method we could use...

https://acwi.gov/monitoring/conference/2008/presentations/sessionE/E6C_Helsel.pdf
https://www.itrcweb.org/gsmc-1/Content/GW%20Stats/5%20Methods%20in%20indiv%20Topics/5%207%20Nondetects.htm



---

clean up code to work with only 1 data frame for comprehension
combining mutates

FOCUS ON WRITING, WRITE ABOUT DIFFERENT METHODS AND UNDERSTANDING THEM BY MONDAY






